# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

The data for this assignment can be found at the following url as a zip file containing the **activity.csv** file in csv file format:

- Url: *https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip*

- File: *activity.csv* 

The variables included in this dataset are:

- **steps**: Number of steps taking in a 5-minute interval (missing
    values are coded as 'NA')

- **date**: The date on which the measurement was taken in YYYY-MM-DD
    format

- **interval**: Identifier for the 5-minute interval in which
    measurement was taken
    

### Loading the data
The following code will be used to download and read the data into the `activity`
data frame:

```{r}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp, method="curl")
activity <- read.csv(unz(temp, "activity.csv"))
unlink(temp)
```

### Analyzing missing values

Exploring the data loaded with the `summary` function:

```{r}
summary(activity)
```

We can see that only the **steps** variable has missing values. The other variables are complete.

Let's analyze how the **steps** variable missing data is distributed by date: 

```{r}
table(activity$date, is.na(activity$steps))
```

We can see that we have 288 measurements each day, since there are 12 x 5 minutes intervals in an hour and 288 x 5 minutes intervals in a day.  
From a total of 61 days with steps measurements, we have data missing from 8 whole days. The other days have complete steps measurements.  
The days that do not have measurements are:

```{r}
unique(activity[is.na(activity$steps),2])
```


### Analyzing data contents

Let's take a look at the structure and the first few rows in the data loaded:

```{r}
str(activity)

head(activity)
```

We notice that:

- the **date** variable contains the data in character format

- the **interval** variable contains numeric value representing the 5 minutes interval the number os steps taken was measured. 

### Preprocessing the data

In order to be able to work with date and time functions to aggregate and plot the data, a new variable, calles `datetime` will be created in the `activity` data frame with the date & time of each of the 5 minutes intervals, in the POSIXct format:

```{r}
activity$datetime<-as.POSIXct(strptime(paste(activity$date, formatC(activity$interval, width=4, flag="0")), "%Y-%m-%d %H%M"))
```


## What is mean total number of steps taken per day?

In order to answer this question, we will create a new variable, called `stepsperday`, with the sum of the number of steps taken each day, being careful to:  
- Use the correct time zone (defult is UTC), and   
- Keep missing values so we can compute the mean correctly with the `mean` function with the parameter `na.rm=TRUE`.  

```{r}
stepsperday<-aggregate(activity$steps, by=list(as.Date(activity$datetime, tz="")), FUN=sum, na.rm=F)

names(stepsperday)<-c("Date", "Steps")

stepsperday
```

So, the mean total number of steps taken per day is:

```{r}
mean(stepsperday$Steps, na.rm=TRUE)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
